---
title: "From Hashtags To Ballots - Analysis of Survey Data"
format: html
editor: visual
---

This document includes the analysis of our survey of German citizens.

## Setup

We load necessary libraries and the main dataset, which already includes the responses of all participants from our quota sample:

```{r}
library(tidyverse)
library(sjmisc)
library(scales)
library(sjPlot)
library(sjlabelled)
library(lubridate)
library(hrbrthemes)
library(cowplot)

df <- read_stata("btw2021_influencers_survey.dta") 
nrow(df)

# removing speeders and incomplete observations
df <- df |>   filter(!is.na(duration) & duration >= 120)
nrow(df)
```

We recode several variables and create new indicators:

```{r}
df <- df |>
  # respondent id
  mutate(respondent_id = str_c("r_", as.character(respondent_id))) |>
  set_na(na = c(99, "99")) 

df <- df |>
  mutate(
    # education as categorical
    edu_cat =
      case_when(
        edu %in% 1:3 ~ 1,
        edu == 4 ~ 2,
        edu %in% 5:6 ~ 3,
        str_contains(edu_other, "Berufsschule") ~ 1,
        str_contains(edu_other, c("Studi|Promotion|Meister")) ~ 3
      ) |> set_labels(labels = c(
        "low" = 1,
        "mid" = 2,
        "high" = 3
      ))
    |> as_factor(),
    # dummy for female participants
    female_dummy = rec(sex, rec = "2=1 [weiblich]; 1=0 [männlich]") |>
      set_label("dummy: frau"),
    # dummy for awareness about influencers
    influ_dummy = rec(influ_aware, rec = "1=1 [ja] ; 2=0 [nein]"),
    # dummy for following influencers
    influ_follow_dummy =  rec(influ_follow, rec = "1=1 [ja] ; 2=0 [nein]"),
    # dummy for low versus high political interest
    polint_dummy = rec(polint, 
           rec = "1:3=0 [wenig pol. Interesse]; 4:5 = 1 [viel pol. Interesse]"),
    polint = as_numeric(polint)
  )

# count indicator for number of social media platforms used
df <- df |> select(why_use_sm01:why_use_sm05)   |>
  as_numeric() |>
  add_columns(df)
df <- df |> select(sm_whatsapp:sm_twitch) |>
  mutate(sm_other = if_else(str_length(sm_other) > 1, "Y", "N")) |>
  row_count(count = "Y",
            var = "sm_count",
            append = FALSE) |>
  set_label("number of social media platforms used") |>
  add_columns(df)

# labels for "helpful for voting decision" indicators
df$voteinfo_mat1 <- set_label(df$voteinfo_mat1, 
                              "election advertizing of parties")
df$voteinfo_mat2 <- set_label(df$voteinfo_mat2,
                              "television news")
df$voteinfo_mat3 <- set_label(df$voteinfo_mat3, 
                              "reports in newspapers & magazines")
df$voteinfo_mat4 <- set_label(df$voteinfo_mat4, 
                              "online information")
df$voteinfo_mat5 <- set_label(df$voteinfo_mat5, 
                              "content of influencers")
df$voteinfo_mat6 <- set_label(df$voteinfo_mat6, 
                              "conversations with relatives & friends")

# creating age / sex table for comparison to German census
df <- df |> mutate(
  age_sex =  case_when(
    female_dummy == 1 & age %in% c(18:29) ~ "Female, Age 18-29",
        female_dummy == 0 & age %in% c(18:29) ~ "Male, Age 18-29",
    female_dummy == 1 & age %in% c(30:39) ~ "Female, Age 30-39",
        female_dummy == 0 & age %in% c(30:39) ~ "Male, Age 30-39",
    female_dummy == 1 & age %in% c(40:49) ~ "Female, Age 40-49",
        female_dummy == 0 & age %in% c(40:49) ~ "Male, Age 40-49",
    female_dummy == 1 & age %in% c(50:59) ~ "Female, Age 50-59",
        female_dummy == 0 & age %in% c(50:59) ~ "Male, Age 50-59",
    female_dummy == 1 & age %in% c(60:99) ~ "Female, Age 60+",
    female_dummy == 0 & age %in% c(60:99) ~ "Male, Age 60+",
    TRUE ~ NA
  ) |> as_factor()
)

# categorical age indicator
df <- df |> mutate(
  age_cat =  case_when(
    age %in% c(18:29) ~ "18-29",
    age %in% c(30:39) ~ "30-39",
    age %in% c(40:49) ~ "40-49",
    age %in% c(50:59) ~ "50-59",
    age %in% c(60:69) ~ "60-69",
    age %in% c(70:99) ~ "70+",
    TRUE ~ NA
  ) |> as_factor()
)
```

## Analysis for Main Paper

### Visualization: Helpfulness for Voting Decision

```{r fig.height=6, fig.width=9}
likert_fig <-  
df |> select(contains("voteinfo_mat")) |> 
  set_labels(labels =  c(
  "1 - not helpful" = "1", 
  "4 - neutral" = "4",
  "7 - very helpful"  =  "7"
))  |>
  plot_likert(
    value = "sum.inside",
    cat.neutral = 4,
    show.n = FALSE,
    legend.title = "Likert scale",
    
    geom.colors = viridis_pal(alpha = 0.8)(6)
  ) + labs(title = NULL) +
  theme_light(base_size = 14) +
  theme(
    legend.position = c(0.85, 0.60),
    legend.background = element_blank(),
    legend.box.background = element_rect(colour = "black")
  )

likert_fig
```

### Visualization: Helpfulness of influencer content by age and sex:

```{r}
df_helpful <-  df |> mutate(
 age_cat2 =  case_when(
    age %in% c(18:29) ~ "18-29",age %in% c(30:39) ~ "30-39",
    age %in% c(40:49) ~ "40-49",
    age %in% c(50:99) ~ "50+",
    TRUE ~ NA) |> as_factor(),
 infl_helpful_dummy = 
    case_when(
      is.na(voteinfo_mat5) ~ NA,
      voteinfo_mat5 %in% c(1,2,3,4) ~"Not helpful /\nneutral",
      TRUE ~ "Helpful") |> as_factor(),
  sex = if_else(female_dummy == 1, "Female respondents", "Male respondents") |> 
           as_factor())


boxplot_fig <- 
df_helpful |> select(infl_helpful_dummy, age, sex) |> drop_na() |> 
  ggplot(aes(x = infl_helpful_dummy, y = age)) +
  geom_violin(aes(fill = infl_helpful_dummy, alpha = 0.90), linewidth = 1) +
  geom_boxplot(width = 0.15,
               color = "black",
               fill = "white") +
  scale_fill_viridis_d(direction = -1, end = 0.95) +
  theme_light()   +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 7)) + 
  theme(
    legend.position = "none") + facet_wrap(~sex) +
  labs( x = "Content of influencers: helpful for voting decision",
        y = "Age of respondents")

boxplot_fig
```

Combining both plots in one figure:

```{r, fig.height= 10, fig.width=8}
plot_grid(likert_fig, boxplot_fig, labels = c('A', 'B'), ncol = 1,
          label_size = 18)

# adjust this path as needed
ggsave("output/fig5.png", bg = "white",
       units = "in", dpi = 300,
     width = 8, height = 10)
```

### Regression Models

The following code computes two logistig regression models for whether survey participants are aware of / follow influencers:

```{r}
# recode social media usage variables as numeric
df <-  df |> mutate(
  why_use_sm04 = as.numeric(why_use_sm04),
  why_use_sm05 = as.numeric(why_use_sm05),
  why_use_sm07 = as.numeric(why_use_sm07)
)

# model for influencer awareness
influ_aware_reg <- glm(
  influ_dummy ~ age + edu_cat + female_dummy +
    sm_count +
    why_use_sm04 + why_use_sm05 +  why_use_sm07 +
    polint,
  data = df,
  family = "binomial"
)

# model for following influencers
influ_follow_reg <- glm(
  influ_follow_dummy ~ age + edu_cat + female_dummy +
    sm_count +
    
    why_use_sm04 + why_use_sm05 +   why_use_sm07 +
    polint  ,
  data = df,
  family = "binomial"
)

# regression table
tab_model( influ_aware_reg, influ_follow_reg)
```

We compare and visualize both models using a forest plot:

```{r fig.height=6, fig.width=9}
plot_models(
  influ_aware_reg,
  influ_follow_reg,
  
  vline.color = "black",
  transform = NULL,
  axis.labels = c(
    "Age",
    "Education: mid (ref. = low)" ,
    "Education: high (ref. = low)",
    "Sex: Female",
    "Number of SM used\n>= once per week",
    "Using SM to follow persons of public interest",
    "Using SM to follow companies or brands",
    "Using SM for entertainment",
    "Interest in politics"
  ) |> rev()
)  +
  theme_light(base_size = 14) +
  scale_y_continuous(limits = c(-1, 1.1)) +
  scale_color_viridis_d(
    end = 0.9,
    labels = c(
      "Following any influencers\n(n = 708)",
      "Awareness of influencers\n(n = 924)"
    )
  ) +
  theme(
    legend.position = c(0.175, 0.17),
    legend.box.background = element_rect(colour = "grey50")
  ) +
  
  labs(
    title = NULL,
    x = "Covariates",
    y = "Logits",
    color = "Dependent Variables"
  )

# adjust this path as needed
ggsave("output/fig4.png", bg = "white",
      units = "in", dpi = 300,
    width = 9, height = 6)
```

## Analysis for Supplementary Material

We inspect frequency tables for socio-demographics of survey respondents:

```{r}

frq(df$age_sex)
frq(df$edu_cat)
frq(df$bl)
```

Creating a figure for the age and sex distribution of participants:

```{r}
df |> filter(is.na(female_dummy) == FALSE) |>
  mutate(men_women = if_else(female_dummy == 1, "Female", "Male") |> 
  as_factor())  |>
  ggplot(aes(x = age_cat, fill = men_women)) +  geom_bar()  +
  theme_light(base_size = 14) +
  scale_fill_viridis_d(alpha = 0.8, begin = 0.2, end = 0.8) +
  
  theme(
    legend.position = c(0.21, 0.6),
    plot.caption = element_text(face = "italic"),
    axis.text.y = element_text(size = 12),
    legend.box.background = element_rect(colour = "grey50")
  ) +
  labs(fill = "Sex", x = "Age", y = "Observations")

# adjust this path as needed
ggsave(
  "output/fig_s4.png",
  bg = "white",
  units = "in",
  dpi = 300,
  width = 9,
  height = 6
)
```

The following code creates a figure for a comparison of voting shares between the GLES Post-Election Survey, our own survey, and the actual results of the German federal election:

```{r fig.height=6, fig.width=9, fig.showtext = TRUE}
party_color <- c(
  "SPD" = "#E3000F",
  "CDU/CSU" = "#000000",
  "GRUENE" = "#1AA037",
  "FDP" =  "#FFEF00",
  "AfD" = "#0489DB",
  "LINKE" = "#E663A6"
)
sources <-  c("BTW21: Results", "GLES Post-Election", "Own Survey")
spd <- c(25.7, 23.5, 27.7)
union <- c(24.1, 17.4, 16.6)
gruene <- c(14.8, 15.4, 15.6)
fdp <- c(11.5, 9.8, 13.3)
afd <- c(10.3, 5.5, 12.3)
linke <- c(4.9, 4.0, 7.3)

shares <- tibble(
  vote_share = c(spd, union, gruene, fdp, afd, linke),
  party = c(
    rep(c("SPD"), 3),
    rep(c("CDU/CSU"), 3) ,
    rep(c("GRUENE"), 3),
    rep(c("FDP"), 3),
    rep(c("AfD"), 3),
    rep(c("LINKE"), 3)
  ) |> as_factor(),
  source = rep(sources, 6) |> as_factor()
)

shares |> ggplot(aes(
  x = reorder(party, vote_share),
  y = vote_share,
  fill = party
)) + geom_col() +
  geom_text(aes(label = vote_share),
            hjust = 1.2,
            color = "grey75") +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 5)) +
  facet_wrap(~ source) + coord_flip() +
  scale_fill_manual(values = party_color) +
  labs(y = "Vote Share", x = "Party", fill = "Party") +
  theme_light(base_size = 14) +
  theme(legend.position = "none",
        strip.text.x = element_text(size = 14))

# adjust this path as needed
ggsave(
  "output/fig_s5.png",
  bg = "white",
  units = "in",
  dpi = 300,
  width = 9,
  height = 6
)
```

At last, we create a figure for voting shares differentiating between respondents who considered influencer content as either helpful or not helpful for their voting decision:

```{r fig.height=6, fig.width=9}
party_color <- c(
  "SPD" = "#E3000F",
  "CDU/CSU" = "#000000",
  "DIE GRÜNEN" = "#1AA037",
  "FDP" =  "#FFEF00",
  "AFD" = "#0489DB",
  "DIE LINKE" = "#E663A6",
  "ANDERE PARTEI" = "grey60"
)

grped <- df |> select(btw_party, voteinfo_mat5) |> drop_na() |>
  mutate(
    party_char = as_label(btw_party) |>
      str_to_upper(),
    influ_vote_helpful = if_else(
      voteinfo_mat5 %in% c("5", "6", "7"),
      "Influencers helpful\nfor voting decision",
      "Influencers not helpful\nfor voting decision"
    )
  )  |>
  filter(!party_char  %in% c("UNGÜLTIGE STIMME ABGEGEBEN")) |>
  
  group_by(influ_vote_helpful, party_char) |> summarise(obs = n()) |>
  ungroup() |>
  group_by(influ_vote_helpful) |>
  mutate(perc = prop.table(obs * 100))

ggplot(grped  , # Draw barplot with grouping & stacking
       aes(x = influ_vote_helpful, y = perc, fill = party_char)) +
  geom_bar(stat = "identity",
           position = "stack",
           alpha = 0.8) +
  scale_fill_manual(values = party_color) + theme_light(base_size = 14) +
  scale_y_percent() +
  
  labs(
    fill = "Vote for\nBTW 2021",
    y = "Percent",
    x = NULL,
    caption = "Dichotomized likert scale (1 - not helpful at all to 7 - very helpful): 1-4 = not helpful, 5-7 = helpful"
  ) +
  
  theme(
    axis.title = element_text(size = 14),
    axis.text.x = element_text(size = 14),
    axis.text.y = element_text(size = 14),
    plot.caption = element_text(lineheight = 1.5, face = "italic")
  )

# adjust this path as needed
ggsave(
  "output/fig_s6.png",
  bg = "white",
  units = "in",
  dpi = 300,
  width = 9,
  height = 6
)
```
